Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-28 VMCAI-revised-version/final.tex
effectivity

\begin{abstract}
\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state in $\Bad$ then it can also leave this bad state and reach a safe state in $\Safe$.
We consider six types of resilience (one of them is the home-space property) defined by an upward-closed set or a downward-closed set $\Safe$, and by the existence of a bound on the length of minimal runs starting from a set $\Bad$ and reaching $\Safe$ ($\Bad$ is generally the complementary of $\Safe$). We study the decidability of each type of resilience for many models: Well Behaved Transition Systems (WBTS), WSTS, VASS and other variations of counter machines. 
%
We first show that all resilience problems are undecidable for effective WSTS with strong compatibility and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) for downward-closed sets $\Safe$.

@@

\begin{proof}
Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute the least $n_0$ by iteratively computing 
$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
returning $n$ if that is the case.
Then, because {\sc $n_0$-state-resilience} is decidable, 
checking $\mathop{\uparrow} \post^*(s) 
 \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
and {\sc State-resilience} is decidable. \qed
\end{proof}

@@

\begin{theorem}\label{post srp}
{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS
when
$\Safe=\mathop{\uparrow} \Safe$. 
\end{theorem}

\begin{proof}
Let $B_{\mathop{\uparrow} \post^*(s)}$ 
 be a basis of $\mathop{\uparrow} \post^*(s)$, and $B_\Safe$ a basis of $\Safe$.
By applying Lemma~\ref{Lemma intersection} twice, we obtain
\[ \post^*(s) 
 \subseteq \pred^*(\Safe) \text{ iff } B_{\mathop{\uparrow} \post^*(s)}  
 \subseteq \pred^*(\Safe)\]
Since $B_{\mathop{\uparrow} \post^*(s)}$ is finite and we can compute 
%\alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ? Est-elle chez Ozkan ?} 
a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\mathop{\uparrow} \post^*(s)} 
 \setminus \pred^*(\Safe) = \emptyset$.  \qed
\end{proof}

>>>>

\begin{abstract}
\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state in $\Bad$ then it can also leave this bad state and reach a safe state in $\Safe$.
We consider six types of resilience (one of them is the home-space property) defined by an upward-closed set or a downward-closed set $\Safe$, and by the existence of a bound on the length of minimal runs starting from a set $\Bad$ and reaching $\Safe$ ($\Bad$ is generally the complementary of $\Safe$). We study the decidability of each type of resilience for many infinite-state transitiuon systems: Well Behaved Transition Systems (WBTS), Well Structured Transition Systems (WSTS), VASS and other variations of counter machines. 
%
We first show that all resilience problems are undecidable for effective WSTS with strong compatibility and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) when $\Safe$ are downward-closed sets.

@@

\begin{proof}
\alain{où utilise-t-on la strong compatibility ? et si on ne l'utilise pas, c'est le theo 8...}
Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute the least $n_0$ by iteratively computing 
$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
returning $n$ if that is the case.
Then, because {\sc $n_0$-state-resilience} is decidable, 
checking $\mathop{\uparrow} \post^*(s) 
 \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
and {\sc State-resilience} is decidable. \qed
\end{proof}

@@

\alain{plus fort que prop 6}
\begin{theorem}\label{post srp}
{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS
when
$\Safe=\mathop{\uparrow} \Safe$. 
\end{theorem}

\begin{proof}
Let $\mathscr{S}$ be a $\pred$-basis-effective, $\post^*$-basis-effective WSTS.
%
Since $\mathscr{S}$ is $\post^*$-basis-effective, let $B_{\mathop{\uparrow} \post^*(s)}$ 
 be a finite basis of $\mathop{\uparrow} \post^*(s)$, hence $\mathop{\uparrow} \post^*(s) = \mathop{\uparrow} B_{\mathop{\uparrow} \post^*(s)}$.
%
Let $B_\Safe$ be a finite basis of $\Safe$, hence $\Safe = .\mathop{\uparrow} B_\Safe$.
%
Since $\mathscr{S}$ is $\pred$-basis-effective, we may compute a finite basis $B_{\pred^*(\Safe)}$ of $\pred^*(\Safe)$ from $B_\Safe$.
%
\alain{je pense qu'il y a une erreur, on n'utilise pas le Lemme 2}
%
By applying Lemma~\ref{Lemma intersection} twice,
we obtain that 
\[ \post^*(s) 
 \subseteq \pred^*(\Safe) \text{ iff } B_{\mathop{\uparrow} \post^*(s)}  
 \subseteq \pred^*(\Safe)\]
%
%
\alain{je pense qu'il y a une erreur, je propose plutôt}
%
Recall that $\mathscr{S}$ is state-resilient iff  $\post^*(s) 
 \subseteq \pred^*(\Safe)$ iff 
\[ \mathop{\uparrow} B_{\mathop{\uparrow} \post^*(s)} 
 \subseteq \mathop{\uparrow}  B_{\pred^*(\Safe)} \text{ iff } \forall b \in B_{\mathop{\uparrow} \post^*(s)}  \exists b' \in B_{\pred^*(\Safe)} \text{ such that } b' \leq b \]
%
Since both $B_{\mathop{\uparrow} \post^*(s)}$ and $B_{\pred^*(\Safe)}$ are finite, {\sc State-resilience} is decidable.
%
\alain{ce qui suit est à enlever je l'ai commenté (à vérifier)}
%
%			Since $B_{\mathop{\uparrow} \post^*(s)}$ is finite
%\alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ? Est-elle chez Ozkan ?} 
%			we can check that $B_{\mathop{\uparrow} \post^*(s)} 
%			 \setminus \pred^*(\Safe) = \emptyset$.  \qed
\end{proof}
